---
title: "Win Ratio OW Relative Efficiency"
author: "Scott Zuo"
output: 
    html_document:
      toc: TRUE
      toc_float: true
      theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Overlap Weighting Estimator

```{r}
start_time <- Sys.time()
```



## Non-Composite Endpoint with Ordinal Outcomes (1 column)

- We want to start with only one ordinal outcome and observe how adjusting for covariates performs; after we see how effective the methods are in this case, we can then expand toward multiple ordinal outcomes or multiple outcomes of mixed distributional types.


## Setting outcome levels


```{r}

#vector of outcomes as a factor with ordered levels
outcomes_3lvl <- factor(c("first", "second", "third"), 
                   levels = c("first", "second", "third"), 
                   ordered = TRUE)

#Order: first < second < third, the larger the better

inv_logit <- function(logit) exp(logit) / (1 + exp(logit))
```




## True Win Ratio


```{r}

#true probabilities trt column

true_logodds1_trt <- 1 + 1 * 1 + 0.5
true_logodds2_trt <- 0.05 + 1 * 1 + 0.5

true_prob_2to3_trt <- inv_logit(true_logodds1_trt)

true_prob_3_trt <- inv_logit(true_logodds2_trt)
true_prob_1_trt <- 1 - true_prob_2to3_trt
true_prob_2_trt <- true_prob_2to3_trt - true_prob_3_trt

#c(true_prob_1_trt,true_prob_2_trt,true_prob_3_trt)
#0.07585818 0.09922809 0.82491373

#true probabilities ctrl column

true_logodds1_ctrl <- 1 + 1 * 1
true_logodds2_ctrl <- 0.05 + 1 * 1

true_prob_2to3_ctrl <- inv_logit(true_logodds1_ctrl)

true_prob_3_ctrl <- inv_logit(true_logodds2_ctrl)
true_prob_1_ctrl <- 1 - true_prob_2to3_ctrl
true_prob_2_ctrl <- true_prob_2to3_ctrl - true_prob_3_ctrl

#c(true_prob_1_ctrl,true_prob_2_ctrl,true_prob_3_ctrl)
# 0.1192029 0.1400222 0.7407749


#trt: 0.07585818 0.09922809 0.82491373
#ctrl: 0.1192029 0.1400222 0.7407749

true_trt_wp_1col <- 0.82491373*(0.1192029+0.1400222) + 0.09922809*0.1192029
true_ctrl_wp_1col <- 0.7407749*(0.07585818+0.09922809) + 0.1400222*0.07585818

#True WR
true_wr_1col <- true_trt_wp_1col/true_ctrl_wp_1col

true_wr_1col
#1.608213
```


## Relative Efficiency for the original and OW Weighted Estimator


```{r}

set.seed(1)

re_count <- 15

empirical_var_ori <- numeric(re_count)
empirical_var_OW <- numeric(re_count)

bias_ori <- numeric(re_count)
bias_OW <- numeric(re_count)

mse_ori <- numeric(re_count)
mse_OW <- numeric(re_count)


for (count_temp in 1:re_count){

n_count <- count_temp*20


trt_eff1 <- 0.5
b1_trt <- 1
b1_ctrl <- 1
b01 <- 1
b02 <- 0.05


###########################################

sim_num <- 1000
WR_sim_1col <- numeric(sim_num)

for (sim_count in 1:sim_num){
  
# x1
x1_trt <- rnorm(n_count, mean = 1, sd = 0.1)
x1_ctrl <- rnorm(n_count, mean = 1, sd = 0.1)

# trt outcome column

logodds1_trt <- b01 + b1_trt * x1_trt + trt_eff1
logodds2_trt <- b02 + b1_trt * x1_trt + trt_eff1

## Probability

prob_2to3_trt <- inv_logit(logodds1_trt)
prob_3_trt <- inv_logit(logodds2_trt)
prob_1_trt <- 1 - prob_2to3_trt
prob_2_trt <- prob_2to3_trt - prob_3_trt

# ctrl outcome column

logodds1_ctrl <- b01 + b1_ctrl * x1_ctrl
logodds2_ctrl <- b02 + b1_ctrl * x1_ctrl

## Probability

prob_2to3_ctrl <- inv_logit(logodds1_ctrl)
prob_3_ctrl <- inv_logit(logodds2_ctrl)
prob_1_ctrl <- 1 - prob_2to3_ctrl
prob_2_ctrl <- prob_2to3_ctrl - prob_3_ctrl
  
#generate random outcomes
outcomes_trt <- c()
for (i in 1:n_count) {
  outcomes_trt[i] <- sample(
    outcomes_3lvl, 
    size = 1,
    prob = c(prob_1_trt[i], prob_2_trt[i], prob_3_trt[i])
  )
}



outcomes_ctrl <- c()
for (i in 1:n_count) {
  outcomes_ctrl[i] <- sample(
    outcomes_3lvl, 
    size = 1,
    prob = c(prob_1_ctrl[i], prob_2_ctrl[i], prob_3_ctrl[i])
  )
}


df_trt_1col <- data.frame(outcomes_trt)
df_ctrl_1col <- data.frame(outcomes_ctrl)

trtwin_1col <- matrix(NA, nrow = nrow(df_trt_1col), ncol = nrow(df_ctrl_1col))
ctrlwin_1col <- matrix(NA, nrow = nrow(df_trt_1col), ncol = nrow(df_ctrl_1col))

for (i in 1:nrow(df_trt_1col)){
  for(j in 1:nrow(df_ctrl_1col)){

      if (df_trt_1col[i,1] > df_ctrl_1col[j,1]){
        trtwin_1col[i,j] = 1
        ctrlwin_1col[i,j] = 0
      }
      else if (df_trt_1col[i,1] < df_ctrl_1col[j,1]){
        trtwin_1col[i,j] = 0
        ctrlwin_1col[i,j] = 1
      }

  }
}


trt_winpr_1col <- sum(as.vector(trtwin_1col), 
                      na.rm = T)/(nrow(df_trt_1col)*nrow(df_ctrl_1col))
ctrl_winpr_1col <- sum(as.vector(ctrlwin_1col), 
                       na.rm = T)/(nrow(df_trt_1col)*nrow(df_ctrl_1col))

WR_sim_1col[sim_count] <- trt_winpr_1col/ctrl_winpr_1col

  
}

WR_sim_noninf_1col <- WR_sim_1col[WR_sim_1col!=Inf & WR_sim_1col!=0 &
                                    !is.na(WR_sim_1col)]


empirical_var_ori[count_temp] <- var(WR_sim_noninf_1col)
bias_ori[count_temp] <- mean(WR_sim_noninf_1col) - true_wr_1col
mse_ori[count_temp] <- bias_ori[count_temp]^2 + empirical_var_ori[count_temp]



###########################################

WR_sim_1col_OW <- numeric(sim_num)

for (sim_count in 1:sim_num){

# x1
x1_trt <- rnorm(n_count, mean = 1, sd = 0.1)
x1_ctrl <- rnorm(n_count, mean = 1, sd = 0.1)

# trt outcome column

logodds1_trt <- b01 + b1_trt * x1_trt + trt_eff1
logodds2_trt <- b02 + b1_trt * x1_trt + trt_eff1

## Probability

prob_2to3_trt <- inv_logit(logodds1_trt)
prob_3_trt <- inv_logit(logodds2_trt)
prob_1_trt <- 1 - prob_2to3_trt
prob_2_trt <- prob_2to3_trt - prob_3_trt

# ctrl outcome column

logodds1_ctrl <- b01 + b1_ctrl * x1_ctrl
logodds2_ctrl <- b02 + b1_ctrl * x1_ctrl

## Probability

prob_2to3_ctrl <- inv_logit(logodds1_ctrl)
prob_3_ctrl <- inv_logit(logodds2_ctrl)
prob_1_ctrl <- 1 - prob_2to3_ctrl
prob_2_ctrl <- prob_2to3_ctrl - prob_3_ctrl

#generate random outcomes
outcomes_trt <- c()
for (i in 1:n_count) {
  outcomes_trt[i] <- sample(
    outcomes_3lvl, 
    size = 1,
    prob = c(prob_1_trt[i], prob_2_trt[i], prob_3_trt[i])
  )
}



outcomes_ctrl <- c()
for (i in 1:n_count) {
  outcomes_ctrl[i] <- sample(
    outcomes_3lvl, 
    size = 1,
    prob = c(prob_1_ctrl[i], prob_2_ctrl[i], prob_3_ctrl[i])
  )
}

df_trt_1col <- data.frame(outcomes_trt)
df_ctrl_1col <- data.frame(outcomes_ctrl)

colnames(df_trt_1col) <- "outcomes_comb"
colnames(df_ctrl_1col) <- "outcomes_comb"

df_comb_1col <- rbind(df_trt_1col, df_ctrl_1col)
df_comb_1col$treatment <- c(rep(1,n_count), rep(0,n_count))
df_comb_1col$x_covariate <- c(x1_trt, x1_ctrl)

###PS
PropScore <- glm(treatment~x_covariate, data = df_comb_1col, family=binomial)
pi_func <- fitted(PropScore)

trtwin_1col <- matrix(NA, nrow = nrow(df_trt_1col), ncol = nrow(df_ctrl_1col))
ctrlwin_1col <- matrix(NA, nrow = nrow(df_trt_1col), ncol = nrow(df_ctrl_1col))

for (i in 1:n_count){
  for(j in (n_count+1):(2*n_count)){

      if (df_comb_1col[i,1] > df_comb_1col[j,1]){
        trtwin_1col[i,j-n_count] = ((1-pi_func[i])*pi_func[j])/2
        ctrlwin_1col[i,j-n_count] = 0
      }
      else if (df_comb_1col[i,1] < df_comb_1col[j,1]){
        trtwin_1col[i,j-n_count] = 0
        ctrlwin_1col[i,j-n_count] = ((1-pi_func[i])*pi_func[j])/2
      }

  }
}


trt_winpr_1col <- sum(as.vector(trtwin_1col), na.rm = T)/(choose(2*n_count,2))
ctrl_winpr_1col <- sum(as.vector(ctrlwin_1col), na.rm = T)/(choose(2*n_count,2))

WR_sim_1col_OW[sim_count] <- trt_winpr_1col/ctrl_winpr_1col

  
}

WR_sim_noninf_1col_OW <- WR_sim_1col_OW[WR_sim_1col_OW!=Inf &
                                        WR_sim_1col_OW!=0 &
                                        !is.na(WR_sim_1col_OW)]


empirical_var_OW[count_temp] <- var(WR_sim_noninf_1col_OW)
bias_OW[count_temp] <- mean(WR_sim_noninf_1col_OW) - true_wr_1col
mse_OW[count_temp] <- bias_OW[count_temp]^2 + empirical_var_OW[count_temp]

}


#empirical_var_ori
#empirical_var_OW

#bias_ori
#bias_OW


relative_eff <- empirical_var_ori/empirical_var_OW



relative_eff
```



## Relative Efficiency Visulization


```{r, warning=FALSE}
library(ggplot2)

var_ori_df <- data.frame(empirical_var_ori)
var_OW_df <- data.frame(empirical_var_OW)

colnames(var_ori_df) <- "variance"
colnames(var_OW_df) <- "variance"

sim_df <- rbind(var_ori_df, var_OW_df)
sim_df$mse <- c(mse_ori, mse_OW)
sim_df$bias <- c(bias_ori, bias_OW)
sim_df$estimator <- c(rep("original",re_count), rep("OW",re_count))
sim_df$estimator <- as.factor(sim_df$estimator)
sim_df$samplesize <- c(c(1:re_count)*20, c(1:re_count)*20)

var_plot <- ggplot(sim_df, aes(x=samplesize, y = variance,
                                  color = estimator)) + 
  geom_point() + 
  geom_line(aes(group = estimator)) + 
  scale_color_manual(values = c("original" = "black", "OW" = "red"))+
  ggtitle("Variance comparison from Simulation") +
  xlab("Sample Size") + ylab("Variance") +
  scale_x_continuous(breaks=seq(0,20*re_count,10)) + 
  theme(
    plot.title = element_text(color="red", size=12, face="bold.italic"),
    axis.title.x = element_text(size=12),
    axis.text.x = element_text(size=6),
    axis.title.y = element_text(size=12),
    legend.text = element_text(size=12)
    )

var_plot


mse_plot <- ggplot(sim_df, aes(x=samplesize, y = mse,
                                  color = estimator)) + 
  geom_point() + 
  geom_line(aes(group = estimator)) + 
  scale_color_manual(values = c("original" = "black", "OW" = "red"))+
  ggtitle("MSE comparison from Simulation") +
  xlab("Sample Size") + ylab("MSE") +
  scale_x_continuous(breaks=seq(0,20*re_count,10)) + 
  theme(
    plot.title = element_text(color="red", size=12, face="bold.italic"),
    axis.title.x = element_text(size=12),
    axis.text.x = element_text(size=6),
    axis.title.y = element_text(size=12),
    legend.text = element_text(size=12)
    )



mse_plot



re_df <- data.frame(relative_eff)
re_df$samplesize <- c(1:re_count)*20

re_plot <- ggplot(re_df, aes(x=samplesize, y = relative_eff)) + 
  geom_point(color = "black") + 
  geom_line(color = "black") +
  geom_hline(yintercept=1, linetype="dashed", color = "red") +
  ggtitle("Relative Efficiency from Simulation, OW vs Original") +
  xlab("Sample Size") + ylab("Relative Efficiency") +
  scale_x_continuous(breaks=seq(0,20*re_count,10)) +
  theme(
    plot.title = element_text(color="red", size=12, face="bold.italic"),
    axis.title.x = element_text(size=12),
    axis.text.x = element_text(size=8),
    axis.title.y = element_text(size=12),
    legend.text = element_text(size=12)
    )

re_plot

```





## Run Time

```{r}
end_time <- Sys.time()

run_time <- end_time - start_time
run_time
```






