---
title: "Win Ratio Doubly Robust Relative Efficiency"
author: "Scott Zuo"
output: 
    html_document:
      toc: TRUE
      toc_float: true
      theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Doubly Robust Estimator

```{r}
library(MASS)
library(nnet)
library(VGAM)
start_time <- Sys.time()
```



## Non-Composite Endpoint with Ordinal Outcomes (1 column)

- We want to start with only one ordinal outcome and observe how adjusting for covariates performs; after we see how effective the methods are in this case, we can then expand toward multiple ordinal outcomes or multiple outcomes of mixed distributional types.


## Setting outcome levels


```{r}

#vector of outcomes as a factor with ordered levels
outcomes_3lvl <- factor(c("first", "second", "third"), 
                   levels = c("first", "second", "third"), 
                   ordered = TRUE)

#Order: first < second < third, the larger the better

inv_logit <- function(logit) exp(logit) / (1 + exp(logit))
```




## True Win Ratio


```{r}

#true probabilities trt column

true_logodds1_trt <- 1 + 1 * 1 + 0.5
true_logodds2_trt <- 0.05 + 1 * 1 + 0.5

true_prob_2to3_trt <- inv_logit(true_logodds1_trt)

true_prob_3_trt <- inv_logit(true_logodds2_trt)
true_prob_1_trt <- 1 - true_prob_2to3_trt
true_prob_2_trt <- true_prob_2to3_trt - true_prob_3_trt

#c(true_prob_1_trt,true_prob_2_trt,true_prob_3_trt)
#0.07585818 0.09922809 0.82491373

#true probabilities ctrl column

true_logodds1_ctrl <- 1 + 1 * 1
true_logodds2_ctrl <- 0.05 + 1 * 1

true_prob_2to3_ctrl <- inv_logit(true_logodds1_ctrl)

true_prob_3_ctrl <- inv_logit(true_logodds2_ctrl)
true_prob_1_ctrl <- 1 - true_prob_2to3_ctrl
true_prob_2_ctrl <- true_prob_2to3_ctrl - true_prob_3_ctrl

#c(true_prob_1_ctrl,true_prob_2_ctrl,true_prob_3_ctrl)
# 0.1192029 0.1400222 0.7407749


#trt: 0.07585818 0.09922809 0.82491373
#ctrl: 0.1192029 0.1400222 0.7407749

true_trt_wp_1col <- 0.82491373*(0.1192029+0.1400222) + 0.09922809*0.1192029
true_ctrl_wp_1col <- 0.7407749*(0.07585818+0.09922809) + 0.1400222*0.07585818

#True WR
true_wr_1col <- true_trt_wp_1col/true_ctrl_wp_1col

true_wr_1col
#1.608213
```



## Relative Efficiency for the original and Doubly Robust Estimator


```{r, warning=FALSE}

set.seed(1)

re_count <- 15

empirical_var_ori <- numeric(re_count)
empirical_var_DR <- numeric(re_count)

bias_ori <- numeric(re_count)
bias_DR <- numeric(re_count)

mse_ori <- numeric(re_count)
mse_DR <- numeric(re_count)

WR_sim_val0_ori <- numeric(re_count)
WR_sim_valInf_ori <- numeric(re_count)
WR_sim_valNA_ori <- numeric(re_count)

WR_sim_val0_DR <- numeric(re_count)
WR_sim_valInf_DR <- numeric(re_count)
WR_sim_valNA_DR <- numeric(re_count)


for (count_temp in 1:re_count){

n_count <- count_temp*20


trt_eff1 <- 0.5
b1_trt <- 1
b1_ctrl <- 1
b01 <- 1
b02 <- 0.05


###########################################

sim_num <- 1000
WR_sim_1col <- numeric(sim_num)
WR_sim_1col_DR <- numeric(sim_num)

for (sim_count in 1:sim_num){
  
# x1
x1_trt <- rnorm(n_count, mean = 1, sd = 0.3)
x1_ctrl <- rnorm(n_count, mean = 1, sd = 0.3)

# trt outcome column

logodds1_trt <- b01 + b1_trt * x1_trt + trt_eff1
logodds2_trt <- b02 + b1_trt * x1_trt + trt_eff1

## Probability

prob_2to3_trt <- inv_logit(logodds1_trt)
prob_3_trt <- inv_logit(logodds2_trt)
prob_1_trt <- 1 - prob_2to3_trt
prob_2_trt <- prob_2to3_trt - prob_3_trt

# ctrl outcome column

logodds1_ctrl <- b01 + b1_ctrl * x1_ctrl
logodds2_ctrl <- b02 + b1_ctrl * x1_ctrl

## Probability

prob_2to3_ctrl <- inv_logit(logodds1_ctrl)
prob_3_ctrl <- inv_logit(logodds2_ctrl)
prob_1_ctrl <- 1 - prob_2to3_ctrl
prob_2_ctrl <- prob_2to3_ctrl - prob_3_ctrl
  
#generate random outcomes
outcomes_trt <- c()
for (i in 1:n_count) {
  outcomes_trt[i] <- sample(
    outcomes_3lvl, 
    size = 1,
    prob = c(prob_1_trt[i], prob_2_trt[i], prob_3_trt[i])
  )
}



outcomes_ctrl <- c()
for (i in 1:n_count) {
  outcomes_ctrl[i] <- sample(
    outcomes_3lvl, 
    size = 1,
    prob = c(prob_1_ctrl[i], prob_2_ctrl[i], prob_3_ctrl[i])
  )
}


df_trt_1col <- data.frame(outcomes_trt)
df_ctrl_1col <- data.frame(outcomes_ctrl)

trtwin_1col <- matrix(NA, nrow = nrow(df_trt_1col), ncol = nrow(df_ctrl_1col))
ctrlwin_1col <- matrix(NA, nrow = nrow(df_trt_1col), ncol = nrow(df_ctrl_1col))

for (i in 1:nrow(df_trt_1col)){
  for(j in 1:nrow(df_ctrl_1col)){

      if (df_trt_1col[i,1] > df_ctrl_1col[j,1]){
        trtwin_1col[i,j] = 1
        ctrlwin_1col[i,j] = 0
      }
      else if (df_trt_1col[i,1] < df_ctrl_1col[j,1]){
        trtwin_1col[i,j] = 0
        ctrlwin_1col[i,j] = 1
      }

  }
}


trt_winpr_1col <- sum(as.vector(trtwin_1col), 
                      na.rm = T)/(nrow(df_trt_1col)*nrow(df_ctrl_1col))
ctrl_winpr_1col <- sum(as.vector(ctrlwin_1col), 
                       na.rm = T)/(nrow(df_trt_1col)*nrow(df_ctrl_1col))

WR_sim_1col[sim_count] <- trt_winpr_1col/ctrl_winpr_1col





###############DR######################

df_trt_1col_DR <- data.frame(outcomes_trt)
df_ctrl_1col_DR <- data.frame(outcomes_ctrl)

colnames(df_trt_1col_DR) <- "outcomes_comb"
colnames(df_ctrl_1col_DR) <- "outcomes_comb"

df_comb_1col <- rbind(df_trt_1col_DR, df_ctrl_1col_DR)
df_comb_1col$treatment <- c(rep(1,n_count), rep(0,n_count))
df_comb_1col$x_covariate <- c(x1_trt, x1_ctrl)

###DR mu-function

if((!(1 %in% df_comb_1col[df_comb_1col$treatment==1,]$outcomes_comb) &
     !(2 %in% df_comb_1col[df_comb_1col$treatment==1,]$outcomes_comb)) |
     (!(1 %in% df_comb_1col[df_comb_1col$treatment==0,]$outcomes_comb) &
     !(2 %in% df_comb_1col[df_comb_1col$treatment==0,]$outcomes_comb))){
  break
}
else{
dr_trt <- vglm(factor(outcomes_comb) ~ x_covariate, data = df_comb_1col[df_comb_1col$treatment==1,],
               family = multinomial)
dr_ctrl <- vglm(factor(outcomes_comb) ~ x_covariate, data = df_comb_1col[df_comb_1col$treatment==0,],
               family = multinomial)

cond_prob_trt <- predictvglm(dr_trt, type = "response")
cond_prob_ctrl <- predictvglm(dr_ctrl, type = "response")
}

prob1_0 <-data.frame(rep(0,n_count))
colnames(prob1_0) <- "1"
prob2_0 <-data.frame(rep(0,n_count))
colnames(prob2_0) <- "2"


if(!"1" %in% colnames(cond_prob_trt)){
  cond_prob_trt <- cbind(prob1_0, cond_prob_trt)
}

if(!"1" %in% colnames(cond_prob_ctrl)){
  cond_prob_ctrl <- cbind(prob1_0, cond_prob_ctrl)
}
if(!"2" %in% colnames(cond_prob_trt)){
  col1 <- data.frame(cond_prob_trt[,1])
  colnames(col1) <- "1"
  col3 <- data.frame(cond_prob_trt[,2])
  colnames(col3) <- "3"
  cond_prob_trt <- cbind(col1, prob2_0, col3)
}

if(!"2" %in% colnames(cond_prob_ctrl)){
  col1 <- data.frame(cond_prob_ctrl[,1])
  colnames(col1) <- "1"
  col3 <- data.frame(cond_prob_ctrl[,2])
  colnames(col3) <- "3"
  cond_prob_ctrl <- cbind(col1, prob2_0, col3)
}




###PS
PropScore <- glm(treatment~x_covariate, data = df_comb_1col, family=binomial)
pi_func <- fitted(PropScore)

trtwin_1col_DR <- matrix(NA, nrow = nrow(df_trt_1col_DR), ncol = nrow(df_ctrl_1col_DR))
ctrlwin_1col_DR <- matrix(NA, nrow = nrow(df_trt_1col_DR), ncol = nrow(df_ctrl_1col_DR))



for (i in 1:n_count){
  for(j in (n_count+1):(2*n_count)){

      if (df_comb_1col[i,1] > df_comb_1col[j,1]){
        trtwin_1col_DR[i,j-n_count] = (0.5)*((1/(pi_func[i]*(1-pi_func[j])))-
          ((1/(pi_func[i]*(1-pi_func[j])))-1)*(cond_prob_trt[i,2]*cond_prob_ctrl[j-n_count,1]+
                                               cond_prob_trt[i,3]*cond_prob_ctrl[j-n_count,1]+
                                               cond_prob_trt[i,3]*cond_prob_ctrl[j-n_count,2])+
                                              (cond_prob_ctrl[j-n_count,2]*cond_prob_trt[i,1]+
                                               cond_prob_ctrl[j-n_count,3]*cond_prob_trt[i,1]+
                                               cond_prob_ctrl[j-n_count,3]*cond_prob_trt[i,2]))
        ctrlwin_1col_DR[i,j-n_count] = (0.5)*(0 -
          ((1/(pi_func[i]*(1-pi_func[j])))-1)*(cond_prob_trt[i,1]*cond_prob_ctrl[j-n_count,2]+
                                               cond_prob_trt[i,1]*cond_prob_ctrl[j-n_count,3]+
                                               cond_prob_trt[i,2]*cond_prob_ctrl[j-n_count,3])+
                                              (cond_prob_ctrl[j-n_count,1]*cond_prob_trt[i,2]+
                                               cond_prob_ctrl[j-n_count,1]*cond_prob_trt[i,3]+
                                               cond_prob_ctrl[j-n_count,2]*cond_prob_trt[i,3]))
      }
      else if (df_comb_1col[i,1] < df_comb_1col[j,1]){
        trtwin_1col_DR[i,j-n_count] = (0.5)*(0 -
          ((1/(pi_func[i]*(1-pi_func[j])))-1)*(cond_prob_trt[i,2]*cond_prob_ctrl[j-n_count,1]+
                                               cond_prob_trt[i,3]*cond_prob_ctrl[j-n_count,1]+
                                               cond_prob_trt[i,3]*cond_prob_ctrl[j-n_count,2])+
                                              (cond_prob_ctrl[j-n_count,2]*cond_prob_trt[i,1]+
                                               cond_prob_ctrl[j-n_count,3]*cond_prob_trt[i,1]+
                                               cond_prob_ctrl[j-n_count,3]*cond_prob_trt[i,2]))
        ctrlwin_1col_DR[i,j-n_count] = (0.5)*((1/(pi_func[i]*(1-pi_func[j])))-
          ((1/(pi_func[i]*(1-pi_func[j])))-1)*(cond_prob_trt[i,1]*cond_prob_ctrl[j-n_count,2]+
                                               cond_prob_trt[i,1]*cond_prob_ctrl[j-n_count,3]+
                                               cond_prob_trt[i,2]*cond_prob_ctrl[j-n_count,3])+
                                              (cond_prob_ctrl[j-n_count,1]*cond_prob_trt[i,2]+
                                               cond_prob_ctrl[j-n_count,1]*cond_prob_trt[i,3]+
                                               cond_prob_ctrl[j-n_count,2]*cond_prob_trt[i,3]))
      }

  }
}


trtwin_1col_DR_2 <-
      matrix(0,
             nrow = nrow(df_trt_1col_DR),
             ncol = nrow(df_ctrl_1col_DR))
ctrlwin_1col_DR_2 <-
      matrix(0,
             nrow = nrow(df_trt_1col_DR),
             ncol = nrow(df_ctrl_1col_DR))
    
for(i2 in 1:n_count-1){
      for (j2 in (i2+1):n_count){
        trtwin_1col_DR_2[i2,j2] <- 0.5*(
          (cond_prob_trt[i2, 2] * cond_prob_trt[j2, 1] +
           cond_prob_trt[i2, 3] * cond_prob_trt[j2, 1] +
           cond_prob_trt[i2, 3] * cond_prob_trt[j2, 2]) + 
          (cond_prob_trt[j2, 2] * cond_prob_trt[i2, 1] +
           cond_prob_trt[j2, 3] * cond_prob_trt[i2, 1] +
           cond_prob_trt[j2, 3] * cond_prob_trt[i2, 2])
          )
        ctrlwin_1col_DR_2[i2,j2] <- 0.5*(
          (cond_prob_ctrl[i2, 2] * cond_prob_ctrl[j2, 1] +
           cond_prob_ctrl[i2, 3] * cond_prob_ctrl[j2, 1] +
           cond_prob_ctrl[i2, 3] * cond_prob_ctrl[j2, 2]) + 
          (cond_prob_ctrl[j2, 2] * cond_prob_ctrl[i2, 1] +
           cond_prob_ctrl[j2, 3] * cond_prob_ctrl[i2, 1] +
           cond_prob_ctrl[j2, 3] * cond_prob_ctrl[i2, 2])
        )
        
        
        
      }
    }
    
    trt_winpr_1col_DR <-
      (sum(as.vector(trtwin_1col_DR), na.rm = T) + 
         sum(as.vector(trtwin_1col_DR_2), na.rm = T)) / (choose(2 * n_count, 2))
    ctrl_winpr_1col_DR <-
      (sum(as.vector(ctrlwin_1col_DR), na.rm = T) +
         sum(as.vector(ctrlwin_1col_DR_2), na.rm = T)) / (choose(2 * n_count, 2))

WR_sim_1col_DR[sim_count] <- trt_winpr_1col_DR/ctrl_winpr_1col_DR


}


WR_sim_noninf_1col <- WR_sim_1col[WR_sim_1col!=Inf &
                                  WR_sim_1col!=0 &
                                  !is.na(WR_sim_1col)]

empirical_var_ori[count_temp] <- var(WR_sim_noninf_1col)
bias_ori[count_temp] <- mean(WR_sim_noninf_1col) - true_wr_1col
mse_ori[count_temp] <- bias_ori[count_temp]^2 + empirical_var_ori[count_temp]


WR_sim_val0_ori[count_temp] <- length(WR_sim_1col[WR_sim_1col = 0])
WR_sim_valInf_ori[count_temp] <- length(WR_sim_1col[WR_sim_1col = Inf])
WR_sim_valNA_ori[count_temp] <- length(WR_sim_1col[is.na(WR_sim_1col)])

###########################################

WR_sim_noninf_1col_DR <- WR_sim_1col_DR[WR_sim_1col_DR!=Inf & 
                                        WR_sim_1col_DR!=0 & 
                                        !is.na(WR_sim_1col_DR)]

empirical_var_DR[count_temp] <- var(WR_sim_noninf_1col_DR)
bias_DR[count_temp] <- mean(WR_sim_noninf_1col_DR) - true_wr_1col
mse_DR[count_temp] <- bias_DR[count_temp]^2 + empirical_var_DR[count_temp]

WR_sim_val0_DR[count_temp] <- length(WR_sim_1col_DR[WR_sim_1col_DR = 0])
WR_sim_valInf_DR[count_temp] <- length(WR_sim_1col_DR[WR_sim_1col_DR = Inf])
WR_sim_valNA_DR[count_temp] <- length(WR_sim_1col_DR[is.na(WR_sim_1col_DR)])

}





relative_eff <- empirical_var_ori/empirical_var_DR
#relative_eff

save(re_count, relative_eff,
     empirical_var_ori, bias_ori, mse_ori,
     empirical_var_DR, bias_DR, mse_DR,
     WR_sim_val0_ori, WR_sim_valInf_ori, WR_sim_valNA_ori,
     WR_sim_val0_DR, WR_sim_valInf_DR, WR_sim_valNA_DR, file = "DR_RE_Output.RData")
```






## Relative Efficiency Visulization


```{r, warning=FALSE}
library(ggplot2)
#load("DR_RE_Output.RData")

var_ori_df <- data.frame(empirical_var_ori)
var_DR_df <- data.frame(empirical_var_DR)

colnames(var_ori_df) <- "variance"
colnames(var_DR_df) <- "variance"

sim_df <- rbind(var_ori_df, var_DR_df)
sim_df$mse <- c(mse_ori, mse_DR)
sim_df$bias <- c(bias_ori, bias_DR)
sim_df$estimator <- c(rep("original",re_count), rep("DR",re_count))
sim_df$estimator <- as.factor(sim_df$estimator)
sim_df$samplesize <- c(c(1:re_count)*20, c(1:re_count)*20)

var_plot <- ggplot(sim_df, aes(x=samplesize, y = variance,
                                  color = estimator)) + 
  geom_point() + 
  geom_line(aes(group = estimator)) + 
  scale_color_manual(values = c("original" = "black", "DR" = "red"))+
  ggtitle("Variance comparison from Simulation") +
  xlab("Sample Size") + ylab("Variance") +
  scale_x_continuous(breaks=seq(0,20*re_count,50)) + 
  theme(
    plot.title = element_text(color="red", size=12, face="bold.italic"),
    axis.title.x = element_text(size=12),
    axis.text.x = element_text(size=12),
    axis.title.y = element_text(size=12),
    legend.text = element_text(size=12)
    )

var_plot


mse_plot <- ggplot(sim_df, aes(x=samplesize, y = mse,
                                  color = estimator)) + 
  geom_point() + 
  geom_line(aes(group = estimator)) + 
  scale_color_manual(values = c("original" = "black", "DR" = "red"))+
  ggtitle("MSE comparison from Simulation") +
  xlab("Sample Size") + ylab("MSE") +
  scale_x_continuous(breaks=seq(0,20*re_count,50)) + 
  theme(
    plot.title = element_text(color="red", size=12, face="bold.italic"),
    axis.title.x = element_text(size=12),
    axis.text.x = element_text(size=12),
    axis.title.y = element_text(size=12),
    legend.text = element_text(size=12)
    )



mse_plot




bias_plot <- ggplot(sim_df, aes(x = samplesize, y = abs(bias))) + 
  geom_point(aes(color = estimator)) + 
  geom_line(aes(color = estimator)) +
  scale_color_manual(values=c("red", "black")) +
  ggtitle("Win Ratio Bias from Simulation") +
  xlab("Sample Size") + ylab("Bias") +
  scale_x_continuous(breaks=seq(0,20*re_count,50)) +
  theme(
    plot.title = element_text(color="red", size=12, face="bold.italic"),
    axis.title.x = element_text(size=12),
    axis.text.x = element_text(size=12),
    axis.title.y = element_text(size=12),
    legend.text = element_text(size=12)
  )




bias_plot


re_df <- data.frame(relative_eff)
re_df$samplesize <- c(1:re_count)*20

re_plot <- ggplot(re_df, aes(x=samplesize, y = relative_eff)) + 
  geom_point(color = "black") + 
  geom_line(color = "black") +
  geom_hline(yintercept=1, linetype="dashed", color = "red") +
  ggtitle("Relative Efficiency from Simulation, Doubly Robust vs Original") +
  xlab("Sample Size") + ylab("Relative Efficiency") +
  scale_x_continuous(breaks=seq(0,20*re_count,50)) +
  theme(
    plot.title = element_text(color="red", size=12, face="bold.italic"),
    axis.title.x = element_text(size=12),
    axis.text.x = element_text(size=12),
    axis.title.y = element_text(size=12),
    legend.text = element_text(size=12)
    )





re_plot

```







## Run Time

```{r}
end_time <- Sys.time()

run_time <- end_time - start_time

run_time
```











