---
title: "Win Ratio Simple Simulation Adjusted for Covariates"
author: "Scott Zuo"
output: 
    html_document:
      toc: TRUE
      toc_float: true
      theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Inverse Propensity Score Weighted Estimator


## Non-Composite Endpoint with Ordinal Outcomes (1 column)

- We want to start with only one ordinal outcome and observe how adjusting for covariates performs; after we see how effective the methods are in this case, we can then expand toward multiple ordinal outcomes or multiple outcomes of mixed distributional types.


### Setting outcome levels


```{r}
#vector of outcomes as a factor with ordered levels
outcomes_3lvl <- factor(c("first", "second", "third"), 
                   levels = c("first", "second", "third"), 
                   ordered = TRUE)

#Order: first < second < third, the larger the better


#number of observations in each group
n_count <- 100
```



### Generating data from covariates

- $$logodds_1 = ln(\frac{\pi_2+\pi_3}{\pi_1})$$
- $$logodds_2 = ln(\frac{\pi_3}{\pi_1+\pi_2})$$


```{r}

set.seed(1)

inv_logit <- function(logit) exp(logit) / (1 + exp(logit))

x1_trt <- rnorm(n_count, mean = 1, sd = 0.1)
trt_eff1 <- 0.5

b1_trt <- 1


x1_ctrl <- rnorm(n_count, mean = 1, sd = 0.1)

b1_ctrl <- 1



# trt outcome column
b01 <- 1
b02 <- 0.05

logodds1_trt <- b01 + b1_trt * x1_trt + trt_eff1
logodds2_trt <- b02 + b1_trt * x1_trt + trt_eff1



## Probability

prob_2to3_trt <- inv_logit(logodds1_trt)
prob_3_trt <- inv_logit(logodds2_trt)

prob_1_trt <- 1 - prob_2to3_trt
prob_2_trt <- prob_2to3_trt - prob_3_trt



# ctrl outcome column


logodds1_ctrl <- b01 + b1_ctrl * x1_ctrl
logodds2_ctrl <- b02 + b1_ctrl * x1_ctrl



## Probability

prob_2to3_ctrl <- inv_logit(logodds1_ctrl)
prob_3_ctrl <- inv_logit(logodds2_ctrl)

prob_1_ctrl <- 1 - prob_2to3_ctrl
prob_2_ctrl <- prob_2to3_ctrl - prob_3_ctrl



```



### True Win Ratio


```{r}

#true probabilities trt column

true_logodds1_trt <- 1 + 1 * 1 + 0.5
true_logodds2_trt <- 0.05 + 1 * 1 + 0.5

true_prob_2to3_trt <- inv_logit(true_logodds1_trt)

true_prob_3_trt <- inv_logit(true_logodds2_trt)
true_prob_1_trt <- 1 - true_prob_2to3_trt
true_prob_2_trt <- true_prob_2to3_trt - true_prob_3_trt

#c(true_prob_1_trt,true_prob_2_trt,true_prob_3_trt)
#0.07585818 0.09922809 0.82491373

#true probabilities ctrl column

true_logodds1_ctrl <- 1 + 1 * 1
true_logodds2_ctrl <- 0.05 + 1 * 1

true_prob_2to3_ctrl <- inv_logit(true_logodds1_ctrl)

true_prob_3_ctrl <- inv_logit(true_logodds2_ctrl)
true_prob_1_ctrl <- 1 - true_prob_2to3_ctrl
true_prob_2_ctrl <- true_prob_2to3_ctrl - true_prob_3_ctrl

#c(true_prob_1_ctrl,true_prob_2_ctrl,true_prob_3_ctrl)
# 0.1192029 0.1400222 0.7407749


#trt: 0.07585818 0.09922809 0.82491373
#ctrl: 0.1192029 0.1400222 0.7407749

true_trt_wp_1col <- 0.82491373*(0.1192029+0.1400222) + 0.09922809*0.1192029
true_ctrl_wp_1col <- 0.7407749*(0.09922809+0.07585818) + 0.1400222*0.07585818

#True WR
true_wr_1col <- true_trt_wp_1col/true_ctrl_wp_1col

true_wr_1col
#1.608213
```



### Simulation without using IPS Weighted Estimator


```{r}

set.seed(1)

sim_num <- 1000
WR_sim_1col <- numeric(sim_num)

for (sim_count in 1:sim_num){
  
#generate random outcomes
outcomes_trt <- c()
for (i in 1:n_count) {
  outcomes_trt[i] <- sample(
    outcomes_3lvl, 
    size = 1,
    prob = c(prob_1_trt[i], prob_2_trt[i], prob_3_trt[i])
  )
}



outcomes_ctrl <- c()
for (i in 1:n_count) {
  outcomes_ctrl[i] <- sample(
    outcomes_3lvl, 
    size = 1,
    prob = c(prob_1_ctrl[i], prob_2_ctrl[i], prob_3_ctrl[i])
  )
}


df_trt_1col <- data.frame(outcomes_trt)
df_ctrl_1col <- data.frame(outcomes_ctrl)

trtwin_1col <- matrix(NA, nrow = nrow(df_trt_1col), ncol = nrow(df_ctrl_1col))
ctrlwin_1col <- matrix(NA, nrow = nrow(df_trt_1col), ncol = nrow(df_ctrl_1col))

for (i in 1:nrow(df_trt_1col)){
  for(j in 1:nrow(df_ctrl_1col)){

      if (df_trt_1col[i,1] > df_ctrl_1col[j,1]){
        trtwin_1col[i,j] = 1
        ctrlwin_1col[i,j] = 0
      }
      else if (df_trt_1col[i,1] < df_ctrl_1col[j,1]){
        trtwin_1col[i,j] = 0
        ctrlwin_1col[i,j] = 1
      }

  }
}


trt_winpr_1col <- sum(as.vector(trtwin_1col), na.rm = T)/(nrow(df_trt_1col)*nrow(df_ctrl_1col))
ctrl_winpr_1col <- sum(as.vector(ctrlwin_1col), na.rm = T)/(nrow(df_trt_1col)*nrow(df_ctrl_1col))

WR_sim_1col[sim_count] <- trt_winpr_1col/ctrl_winpr_1col

  
}

WR_sim_noninf_1col <- WR_sim_1col[WR_sim_1col!=Inf & WR_sim_1col!=0 &
                                    !is.na(WR_sim_1col)]

summary(WR_sim_noninf_1col)
var(WR_sim_noninf_1col)


```




### Simulation using IPS Weighted Estimator

```{r}
set.seed(1)

sim_num <- 1000
WR_sim_1col_IPS <- numeric(sim_num)

for (sim_count in 1:sim_num){
  
#generate random outcomes
outcomes_trt <- c()
for (i in 1:n_count) {
  outcomes_trt[i] <- sample(
    outcomes_3lvl, 
    size = 1,
    prob = c(prob_1_trt[i], prob_2_trt[i], prob_3_trt[i])
  )
}



outcomes_ctrl <- c()
for (i in 1:n_count) {
  outcomes_ctrl[i] <- sample(
    outcomes_3lvl, 
    size = 1,
    prob = c(prob_1_ctrl[i], prob_2_ctrl[i], prob_3_ctrl[i])
  )
}


df_trt_1col <- data.frame(outcomes_trt)
df_ctrl_1col <- data.frame(outcomes_ctrl)

colnames(df_trt_1col) <- "outcomes_comb"
colnames(df_ctrl_1col) <- "outcomes_comb"

df_comb_1col <- rbind(df_trt_1col, df_ctrl_1col)
df_comb_1col$treatment <- c(rep(1,n_count), rep(0,n_count))
df_comb_1col$x_covariate <- c(x1_trt, x1_ctrl)

###IPS
PropScore <- glm(treatment~x_covariate, data = df_comb_1col, family=binomial)
pi_func <- fitted(PropScore)

trtwin_1col <- matrix(NA, nrow = nrow(df_trt_1col), ncol = nrow(df_ctrl_1col))
ctrlwin_1col <- matrix(NA, nrow = nrow(df_trt_1col), ncol = nrow(df_ctrl_1col))

for (i in 1:n_count){
  for(j in (n_count+1):(2*n_count)){

      if (df_comb_1col[i,1] > df_comb_1col[j,1]){
        trtwin_1col[i,j-n_count] = (1/(pi_func[i]*(1-pi_func[j])))/2
        ctrlwin_1col[i,j-n_count] = 0
      }
      else if (df_comb_1col[i,1] < df_comb_1col[j,1]){
        trtwin_1col[i,j-n_count] = 0
        ctrlwin_1col[i,j-n_count] = (1/(pi_func[i]*(1-pi_func[j])))/2
      }

  }
}


trt_winpr_1col <- sum(as.vector(trtwin_1col), na.rm = T)/(choose(2*n_count,2))
ctrl_winpr_1col <- sum(as.vector(ctrlwin_1col), na.rm = T)/(choose(2*n_count,2))

WR_sim_1col_IPS[sim_count] <- trt_winpr_1col/ctrl_winpr_1col

  
}

WR_sim_noninf_1col_IPS <- WR_sim_1col_IPS[WR_sim_1col_IPS!=Inf & WR_sim_1col_IPS!=0 &
                                            !is.na(WR_sim_1col_IPS)]

summary(WR_sim_noninf_1col_IPS)
var(WR_sim_noninf_1col_IPS)

```

- The way we set up the code:
  - will only include the non-zero values among all comparisons
  - $W_i = 1$ and $W_j = 0$ should be guaranteed



### Simulation results display and comparison (one ordinal outcome)


```{r, warning=FALSE}
library(ggplot2)

WR_sim_df_1col <- data.frame(WR_sim_noninf_1col)
WR_sim_df_1col_IPS <- data.frame(WR_sim_noninf_1col_IPS)


line.data_1col <- data.frame(xintercept = c(mean(WR_sim_noninf_1col), median(WR_sim_noninf_1col), true_wr_1col), 
                        Lines = c("Mean", "Median", "True WinRatio"),
  color = c("black", "red", "gold"), stringsAsFactors = FALSE)

line.data_1col_IPS <- data.frame(xintercept = c(mean(WR_sim_noninf_1col_IPS), median(WR_sim_noninf_1col_IPS), 
                                                true_wr_1col), 
                        Lines = c("Mean", "Median", "True WinRatio"),
  color = c("black", "red", "gold"), stringsAsFactors = FALSE)


WR_hist1_1col <- ggplot(WR_sim_df_1col, aes(x=WR_sim_noninf_1col)) + 
  geom_histogram(color="grey", fill="white", binwidth = 1, aes(y=after_stat(density))) +
  xlim(0, 10) +
  geom_density(alpha=0.2, color = "grey", fill="salmon") +
  geom_vline(aes(xintercept = xintercept, color = Lines), line.data_1col, linewidth = 0.6) +
  scale_colour_manual(values = line.data_1col$color) +
  ggtitle("Simulation without adjusting for covariates") +
  xlab("Win Ratio value") +
  theme(
    plot.title = element_text(color="red", size=12, face="bold.italic"),
    axis.title.x = element_text(size=12),
    axis.title.y = element_text(size=12),
    legend.text = element_text(size=12)
    )

WR_hist1_1col_IPS <- ggplot(WR_sim_df_1col_IPS, aes(x=WR_sim_noninf_1col_IPS)) + 
  geom_histogram(color="grey", fill="white", binwidth = 1, aes(y=after_stat(density))) +
  xlim(0, 10) +
  geom_density(alpha=0.2, color = "grey", fill="salmon") +
  geom_vline(aes(xintercept = xintercept, color = Lines), line.data_1col_IPS, linewidth = 0.6) +
  scale_colour_manual(values = line.data_1col_IPS$color) +
  ggtitle("Simulation using Inverse Propensity Score Weighted Estimator") +
  xlab("Win Ratio value") +
  theme(
    plot.title = element_text(color="red", size=12, face="bold.italic"),
    axis.title.x = element_text(size=12),
    axis.title.y = element_text(size=12),
    legend.text = element_text(size=12)
    )

WR_hist1_1col
WR_hist1_1col_IPS
```





## Composite Endpoint with Ordinal Outcomes (more than 1 column)

- In the setting of composite endpoint, we will analyze the Win Ratio with two or more columns of ordinal outcomes in each group









